---
title: "Class 11: Structural Bioinformatics pt2"
author: "Mario Gutierrez (PID: A69043009)"
format: pdf

---

## Bacground

We saw last that the PDB has 209,886 entries (Oct/Nov 2025). UniProtKB (i.e. protein sequence database) has 199,579,901 entries.

```{r}
209886/199579901 * 100
```

So the PDB has only 0.1% coverage of the main sequence database.

Enter AlphaFold data base (AFDB) < https://alphafold.ebi.ac.uk > that attempts to provide computed models for all sequences in UniProt. 

"AlphaFold DB provides open access to over 200 million protein structure predictions to accelerate scientific research."

## AlphaFold

AlphaFold has 3 main outputs

- the predicted coordinates (PDB files)
- A local quality score called **pLDDT** (one for each amino acid)
- A second quality score **PAE** Predicted Aligned Error (for each pair of amino acid)


We can run AlphaFold ourselves if we are not happy with AFDB (i.e. no coverage or poor model).


## Interpreting/analyzing AF results in R

```{r}
results_dir <- "HIVPR_dimer_23119/"
```

```{r}
pdb_files <- list.files(path=results_dir,
                        pattern="*.pdb",
                        full.names = TRUE)

# Print our PDB file names
basename(pdb_files)
```

```{r}
library(bio3d)
library(bio3d.view)
pdbs <- pdbaln(pdb_files, fit=TRUE, exefile="msa")
```


```{r}
#library(rgl)
library(bio3d)
library(bio3d.view)


view.pdbs(pdbs)
```


```{r}
aln_file <- list.files(path=results_dir,
                       pattern=".a3m$",
                        full.names = TRUE)
aln_file

```


```{r}
aln <- read.fasta(aln_file[1], to.upper = TRUE)
```

```{r}
sim <- conserv(aln)
```







