---
title: "Class 08: Breast Cancer Analysis Project"
author: "Mario Gutierrez (PID: A69043009)"
format: pdf
toc: true
---

## Backround

The goal of this mini-project is  to explore a complete analysis using the unsupervised learning techniques covered in class. We will extend what you’ve learned by combining PCA as a preprocessing step to clustering using data that consist of measurements of cell nuclei of human breast masses. 

The data itself comes from the Wisconsin Breast Cancer Diagnostic Data Set first reported by K. P. Benne and O. L. Mangasarian: “Robust Linear Programming Discrimination of Two Linearly Inseparable Sets”.

Values in this data set describe characteristics of the cell nuclei present in digitized images of a fine needle aspiration (FNA) of a breast mass.

## Data import

The data is available as a CSV from the class website:
```{r}
wisc.df <- read.csv("WisconsinCancer.csv", row.names = 1)
```

Make sure we do not include sample id or diagnosis columns in the data that we analyze below.

```{r}
diagnosis <- as.factor(wisc.df$diagnosis)
wisc.data <- wisc.df[,-1]
dim(wisc.data)
```
```{r}
head(wisc.data)
```

> Q1. How many observations are in this dataset?

There are `r nrow(wisc.data)` observations.

```{r}
nrow(wisc.data)
```

> Q2. How many of the observations have a malignant diagnosis?

```{r}
sum(wisc.df$diagnosis=="M")
```

```{r}
table(wisc.df$diagnosis)
```

> Q3. How many variables/features in the data are suffixed with _mean?

```{r}
#colnames(wisc.data)

length(grep("_mean", colnames(wisc.data)))

```
```{r}
n <- colnames(wisc.data)
inds <- grep("_mean", n)
length(inds)
```
## Principal Component Analysis

The main function in base R for PCA is called `prcomp()`. An optional argument `scale` should nearly always be switched to `scale=TRUE` for this function.

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```
Let's make our main result figure - the "PC plot" or "score plot", " ordienation plot"...

```{r}
library(ggplot2)
ggplot(wisc.pr$x) +
  aes(PC1, PC2, col=diagnosis) +
  geom_point()


```

> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

44%

```{r}
pr.var <- wisc.pr$sdev^2
pr.var/sum(pr.var)
```


> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

PC1 and PC2

> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

Up until PC6

## Hierarchical clustering
```{r}
d <- dist(scale(wisc.data))
h <- hclust(d)
plot(h)
```


```{r}
dist_matrix <- dist(wisc.pr$x[, 1:7])
wisc.pr.hclust <- hclust(d = dist_matrix, "ward.D2")
plot(wisc.pr.hclust)
```

## Combining PCA and clustering

```{r}
d <- dist(wisc.pr$x[,1:3])
wisc.pr.hclust <- hclust(d, method = "ward.D2")
plot(wisc.pr.hclust)
abline(h=60, col="red")
```

Get my cluster membership vector
```{r}
grps <- cutree(wisc.pr.hclust, h=70)
table(grps)
```

```{r}
table(diagnosis)
```

Make a wee "cross-table"

```{r}
table(grps, diagnosis)
```

TP: 179
FP: 24

Sensitivity: TP/(TP+FN)

> Q14. How well do the hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.


```{r}
wisc.pr.hclust.clusters <- cutree(h, k=2)
table(wisc.pr.hclust.clusters, diagnosis)
```

Hierarchical clustering previous to PCA makes B and M samples group together in one single cluster, having too many false negatives. Therefore, the combined method works better.




