---
title: "Class05: Data Viz with GGPLOT"
author: "Mario Gutierrez (PID A69043009)"
format: pdf
editor: visual
---

Today we are playing with plotting and graphics in R.

There are lots of ways to make cool figures in R.
There is "base" R graphics (`plot()`, `hist()`, `boxplot()`, etc.)

There is also add-on packages, like **ggplot**

```{r}
head(cars, 3)
```
Let's plot this with "base" R:

```{r}
plot(cars)
```
```{r}
head(mtcars)
```
Let's plot `mpg` vs `disp`
```{r}
plot(mtcars$mpg, mtcars$disp, pch=10)
```
```{r}
hist(mtcars$mpg)
```

## GGPLOT

The main function in the ggplot2 package is **ggplot**. First I need to install the **ggplot2** package. I can install any package with the function `install.packages()`

> **N.B.** I never want to run `install.packages` in my quarto source document!!!

```{r}
library(ggplot2)
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point()
```
```{r}
ggplot(cars) +
  aes(speed) +
  geom_histogram()
```

Every ggplot needs at least 3 things:

- The **data** (given with `ggplot(cars)`)
- The **aesthetic** mapping (given with `aes()`)
- The **geom** (given by `geom_point()`)

> For simple canned plots or graphs "base" R is nearly always faster.

### Adding more layers

Let's add a line

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  geom_line()
```

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE)
```

Let's add a line and a title, subtitle and caption as well as custom axis labels.

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Silly Plot",
       x="Speed (MPH)",
       y= "Distance (ft)",
       caption="when is tea time anyway???") +
  theme_bw()
```

## Plot some expression data

Read file on line
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
> Q1. How many genes are in this wee daaset?

There are `r nrow(genes)` in this dataset

> Q2. How may "up" regulated genes are there?

```{r}
sum(genes$State == "up")
```
There are `r sum(genes$State == "up")` of genes
```{r}
table(genes$State)
```

```{r}
ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=Condition2) +
  geom_point()
```
```{r}
p <- ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point() +
  scale_color_manual(values=c("blue", "black", "red"))


```

```{r}
p 
```
```{r}
```{r}
ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=State,
      label=Gene) +
  geom_point() +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_text()


```


```{r}
library(ggrepel)

ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=State,
      label=Gene) +
  geom_point() +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_text_repel(max.overlaps=60) +
  theme_bw()
```

## Going Further

Playing with some different layers and the gapminder dataset...

```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

```{r}
head(gapminder)
```
```{r}
tail(gapminder)
```
```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, col=continent) +
  geom_point(alpha=0.2) +
  facet_wrap(~continent) +
  theme_bw()
```
```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, col=continent) +
  geom_violin() + geom_point() +
  facet_wrap(~continent) +
  theme_bw()
```


