---
title: "Class 17: Introduction to Genome Informatics Lab"
author: "Mario Gutierrez (PID: A69043009)"
format: pdf
---

## Section 1. Proportion of G/F in a population

Downloaded a CSV file from Ensemble <https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=959672880#373531_tablePanel> 

Here we read this CSV file 

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(mxl)
```

> Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are
homozygous for the asthma associated SNP (G|G)?

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

Now let's look at a different population. I picked the GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(gbr)
```

Find proportion of G|G

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) * 100, 2)
```
This varient that is associated with childhood asthma is more frequent in the GBR population than the MXL population. 

Let's now dig into this further.

## Section 4: Population Scale Analysis

One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 

So, you processed about ~230 samples and did the normalization on a genome level. Now, you want to find whether there is any association of the 4 asthma-associated SNPs (rs8067378…) on ORMDL3 expression.

How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")

head(expr)
```
> Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
median_AA <- median(expr$exp[expr$geno == "A/A"])
median_AG <- median(expr$exp[expr$geno == "A/G"])
median_GG <- median(expr$exp[expr$geno == "G/G"])

median_AA
median_AG
median_GG
```

```{r}
library(ggplot2)

```
> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

Yes, the SNP affects ORMDL3 expression, making the homozygous G|G have lower expression levels compared to the heterozygous A|G an normal homozygous A|A. 

Let's make a boxplot


```{r}
ggplot(expr) +
  aes(x=geno, y=exp, fill=geno) +
  geom_boxplot(notch = TRUE)
```






