---
title: "Class 7: Machine Learning 1"
author: "Mario Gutierrez PID: A69043009"
format: pdf

---

Today we will begin our exploratioon of some "classical" machine learning approaches. We will star with clustering:

Let's first make up some data to cluster where we know what the answer should be.

```{r}
rnorm(1000)

```
```{r}
hist(rnorm(10000))
```
```{r}
x <- c(rnorm(30, mean=-3), rnorm(30, mean = 3))
y <- rev(x)

x <- cbind(x, y)
head(x)
```

```{r}
plot(x)
```

The main function in "base" R for Kmeans clustering is called `kmeans()`.

```{r}
k <- kmeans(x, centers = 2)
k
```

> Q. How big are the clusters (i.e. their size)?

```{r}
k$size
```

> Q. What clusters do my data points reside in?

```{r}
k$cluster
```

> Q. Make a plot of our data colored by cluster assignment - i.e. Make a result figure...

```{r}
plot(x, col=k$cluster)
points(k$centers, col="blue", pch=15)

```

> Q. Cluster with k-means into 4 clusters and plot your results as above

```{r}
k4 <- kmeans(x, centers = 4)
plot(x, col=k4$cluster)
points(k4$centers, col="blue", pch=15)
```
> Q. Run kmeans with centers (i.e. values of k) equal 1 to 6.

```{r}
k1 <- kmeans(x, centers = 1)$tot.withinss
k2 <- kmeans(x, centers = 2)$tot.withinss
k3 <- kmeans(x, centers = 3)$tot.withinss
k4 <- kmeans(x, centers = 4)$tot.withinss
k5 <- kmeans(x, centers = 5)$tot.withinss
k6 <- kmeans(x, centers = 6)$tot.withinss

ans <- c(k1, k2, k3, k4, k5, k6)
```

Use for loop

```{r}
ans <- NULL
for (i in 1:6) {
  ans <- c(ans, kmeans(x, centers = i)$tot.withins)
  
}
ans
```

Make a "scree-plot"
```{r}
plot(ans, type = "b")
```

## Hierarchical Clustering

The main function in "base" R for this is called `hclust()`
```{r}
d <- dist(x)
hc <- hclust(d)
hc
```

```{r}
plot(hc)
abline(h=7, col="red")
```

To obtain clusters from our `hclust` result object **hc* we "cut" the tree to yield different sub branches. For this we use `cutree()` function.

```{r}
grps <- cutree(hc, h=7)
grps
```
```{r}
plot(x, col=grps)
```

```{r}
library(pheatmap)
pheatmap(x)
```

## Principal Component Analysis (PCA)

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```

>Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
dim(x)
```

```{r}
rownames(x) <- x[,1]
x <- x[,-1]

```


```{r}
x <- read.csv(url, row.names=1)
head(x)
```

> Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

I like fixing it up fron when reading my data

## Spotting major differences and trends

```{r}
rainbow(nrow(x))
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

```{r}
rainbow(nrow(x))
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```



## Pair and heatmaps

Scatterplot matrices )a.k.a. "pairs plots) can be useful for relativity samll datasets like this one. Let's have a look:

```{r}
pairs(x, col=rainbow(nrow(x)), pch=16)

```

```{r}
library(pheatmap)

pheatmap( as.matrix(x) )
```

> Q6. Based on the pairs and heatmap figures, which countries cluster together and what does this suggest about their food consumption patterns? Can you easily tell what the main differences between N. Ireland and the other countries of the UK in terms of this data-set?

Wales and England are similar in consumption of these foods. It is still quite difficult to say what is going on in the dataset. 


## PCA to the rescue

The main function in "base" R for PCA is called `prcomp()`.

As we want to do PCA on the food data for the different countries we will want the foods in the columns.

```{r}
pca <- prcomp(t(x))
summary(pca)
```

Our result object is called `pca` and it has a `$x` component that we will look at first

```{r}
pca$x
```

```{r}
library(ggplot2)
cols <- c("orange", "red", "blue", "green")

ggplot(pca$x) +
  aes(PC1, PC2, label=rownames(pca$x)) +
  geom_point()+
  geom_text()
```

Another major result out of PCA is the so-called "variable loadings" or `$rotation` that tells us how the original viariables (foods) contribute to PCs (i.e. our new axis). 

```{r}
ggplot(pca$rotation)+
  aes(PC1, rownames(pca$rotation))+
  geom_col()

```




