---
title: "Class 16: Analyzing sequencing data in the cloud"
author: "Mario Gutierrez (PID: A69043009)"
format: pdf
toc: true
---

##  Downstream analysis

```{r}
library(tximport)

# setup the folder and filenames to read
folders <- dir(pattern="SRR21568*")
samples <- sub("_quant", "", folders)
files <- file.path( folders, "abundance.h5" )
names(files) <- samples

txi.kallisto <- tximport(files, type = "kallisto", txOut = TRUE)
```
```{r}
head(txi.kallisto$counts)
```

Number of transcripts each sample has:
```{r}
colSums(txi.kallisto$counts)
```

Number of transcripts detected in at least one sample
```{r}
sum(rowSums(txi.kallisto$counts)>0)
```

Filter out annotated transcripts with no reads:
```{r}
to.keep <- rowSums(txi.kallisto$counts) > 0
kset.nonzero <- txi.kallisto$counts[to.keep,]
```

Filter out transcripts with no change over the sample:

```{r}
keep2 <- apply(kset.nonzero,1,sd)>0
x <- kset.nonzero[keep2,]
```


## Principal components analysis

We compute the PCA scaling the transcripts levels so that each feature contribute equally to the PCA.

```{r}
pca <- prcomp(t(x), scale=TRUE)
```

```{r}
summary(pca)
```

With the PCA we can visualize the summarized transcriptomic profiles of each sample:

```{r}
plot(pca$x[,1], pca$x[,2],
     col=c("blue","blue","red","red"),
     xlab="PC1", ylab="PC2", pch=16)
```

Using ggplot for the same PCA:
PC1 vs PC2

```{r}
library(ggplot2)

df <- data.frame(
  PC1 = pca$x[,1],
  PC2 = pca$x[,2],
  group = factor(c("blue","blue","red","red"))
)

ggplot(df) + 
  aes(x = PC1, y = PC2, color = group) +
  geom_point(size = 3) +
  scale_color_manual(values = c("blue" = "blue", "red" = "red")) +
  labs(x = "PC1", y = "PC2") +
  theme_bw()

```

PC1 vs PC3

```{r}
df <- data.frame(
  PC1 = pca$x[,1],
  PC3 = pca$x[,3],
  group = factor(c("blue","blue","red","red"))
)

ggplot(df) + 
  aes(x = PC1, y = PC3, color = group) +
  geom_point(size = 3) +
  scale_color_manual(values = c("blue" = "blue", "red" = "red")) +
  labs(x = "PC1", y = "PC3") +
  theme_bw()
```

PC2 vs PC3

```{r}
df <- data.frame(
  PC2 = pca$x[,2],
  PC3 = pca$x[,3],
  group = factor(c("blue","blue","red","red"))
)

ggplot(df) + 
  aes(x = PC2, y = PC3, color = group) +
  geom_point(size = 3) +
  scale_color_manual(values = c("blue" = "blue", "red" = "red")) +
  labs(x = "PC2", y = "PC3") +
  theme_bw()
```



## Differential Expression Analysis
```{r, message=FALSE}
library(DESeq2)
```

```{r}
sampleTable <- data.frame(condition = factor(rep(c("control", "treatment"), each = 2)))
rownames(sampleTable) <- colnames(txi.kallisto$counts)
```

```{r}
dds <- DESeqDataSetFromTximport(txi.kallisto,
                                sampleTable, 
                                ~condition)
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
head(res)
```




