---
title: "Class 10: Halloween Mini-Project"
author: "Mario Gutierrez (PID: A69043009)"
format: pdf

---

## Halloween Mini-Project

## Backround 

It is Halloween today and to help get in the mood for tonight let's explore some candy preference data from 538.


```{r}
candy_file <- read.csv("candy-data.csv")
candy = read.csv("candy-data.csv", row.names=1)
head(candy)
```

```{r}
library(flextable)
flextable::flextable(head(candy))
```

```{r, message=FALSE}
library(dplyr)
nrow(candy)

```


```{r}
round(mean(candy$winpercent), 4)
```

```{r}
candy |> select(winpercent)
```

```{r}
win <- candy$winpercent
win.mean <- mean(win)
round(win.mean)
```

```{r}
candy %>% select(winpercent)


candy |> 
  select(winpercent)
```


> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Milky Way", ]$winpercent
```

```{r}
colnames(candy)
```




```{r}
library(skimr)
skim(candy)
```


> Q8. Plot a histogram of winpercent values

```{r}
library(ggplot2)

ggplot(candy)+
  aes(winpercent)+
  geom_histogram(bins=20)
```

> Q9. Is the distribution of winpercent values symmetrical?

```{r}
ggplot(candy)+
  aes(winpercent)+
  geom_density()
```

> Q10. Is the center of the distribution above or below 50%?

```{r}
mean(candy$winpercent)
```
```{r}
summary(candy$winpercent)
```

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
# 1. Find all chocolate candy in the dataset
choc.ids <- as.logical(candy$chocolate)
choc.candy <- candy[choc.ids,]
choc.candy
# 2. Eextract their `windpercent` values
choc.win <- choc.candy$winpercent

# 3. Find the mean of these values
choc.mean <- mean(choc.win)
#4-6 Do the same for fruity candy
fruit.win <- candy[candy$fruity==1,]$winpercent
fruit.mean <- mean(fruit.win)
# Which mean value is higher
choc.mean
fruit.mean
```



> Q12. Is this difference statistically significant?



```{r}
t.test(choc.win, fruit.win)
```

> Q13. What are the five least liked candy types in this set?

```{r}
ord.ind <- order(candy$winpercent)
head(candy[ord.ind,], 5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[ord.ind,], 5)
```
> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
ggplot(candy)+
  aes(x=winpercent, y=reorder(rownames(candy), winpercent))+
  geom_col()
  
```

```{r}
mycols <- rep("black", nrow(candy))
mycols[candy$chocolate==1] <- "chocolate"
mycols[candy$bar==1] <- "brown"
mycols[candy$fruity==1] <- "pink"
mycols

ggplot(candy)+
  aes(x=winpercent, y=reorder(rownames(candy), winpercent))+
  geom_col(fill=mycols)
  
```


## Winpercent vs Pricepercent

```{r}
library(ggrepel)

mycols[candy$fruity==1] <- "red"

ggplot(candy)+
  aes(winpercent, pricepercent, label=rownames(candy))+
  geom_point(col=mycols)+
  geom_text_repel(col=mycols)
  
```

## Correlation


```{r}
cij <- cor(candy)
```


```{r}
library(corrplot)
corrplot(cij)
```


## Principal Component Analysis

The main funciton in base R for this `prcomp()` and we want to set `scales=TRUE` HERE:


```{r}
pca <- prcomp(candy, scale. = TRUE)
summary(pca)
```

Let's look at our first main result figure - the "PC plot" or PC1 vs PC2

```{r}
ggplot(pca$x)+ 
  aes(PC1, PC2, label=rownames(pca$x))+
  geom_point(col=mycols)+
  geom_text_repel(col=mycols)+
  labs(titlle="Barry's candy ")
```

Don't forget about your variable "loadings" - how the original variables contribute to your new PCs...

```{r}
ggplot(pca$rotation)+
  aes(PC1, rownames(pca$rotation))+
  geom_col()
```





